---
title: 'Final Project: Visualizing Coronavirus Data'
author: 'Name: Sal Figueroa'
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: 3
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r setup}
#### Load necessary library ####
packages <- c("knitr", "kableExtra", "magrittr", "readr", "geosphere")

install_me <- packages[!(packages %in% installed.packages()[, "Package"])]
if (length(install_me)) install.packages(install_me)

library(knitr)
library(magrittr)
library(readr)
library(geosphere)
library(ggplot2)
library(treemapify)
library(tidyverse)
library(ggfittext)
library(scales)
library(dplyr)
library(data.table)
library(ggrepel)
library(scales)
library(stringr)
library(leaflet)
library(kableExtra)
library(gridExtra)
#library(cowplot)

knitr::opts_chunk$set(echo=TRUE)
```



## Github Repository

*Repository holds all related files to Project*

[github: Figgs0bit-Final Project (https://github.com/Figgs0bit/CSIT165-CovidGroupProj-2)](https://github.com/Figgs0bit/CSIT165-CovidGroupProj-2)


## Required data sets

*Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE). Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. Country/region are countries/regions hat conform to World Health Organization (WHO). Lat and Long refer to coordinates references for the user. Date fields are stored in MM/DD/YYYY format.*

*For this project, we will use ALL of the data sets provided in this GitHub repository. These include global data sets for COVID-19 associated confirmations and deaths as well as COVID-19 data sets associated with confirmations and deaths for cities in the US.*



[2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by John Hopkins CSSE](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series)

*-time_series_covid19_deaths_global.csv*

*-time_series_covid19_deaths_US.csv*

*-time_series_covid19_confirmed_global.csv*

*-time_series_covid19_confirmed_US.csv*


### Data Load

```{r Data load}
print("Lab represents data download May 20th, 2025")

#time_series_covid19_confirmed_global.csv, Rows-289 and Cols-1147.
confirmed_global <- as.data.frame(read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/refs/heads/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"))

#time_series_covid19_deaths_global.csv, Rows-289 and Cols-1147.
deaths_global <- as.data.frame(read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/refs/heads/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"))

#time_series_covid19_deaths_global.csv, Rows-3342 and Cols-1155 Has population
deaths_US <- as.data.frame(read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/refs/heads/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"))

#loads the population column from data frame Deaths_US
Population <- deaths_US[12]

#time_series_covid19_confirmed_global.csv, Rows-3342 and Cols-1154
confirmed_US <- as.data.frame(read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/refs/heads/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"))

#Adds Population column to confirmed_us dataframe.
confirmed_US <- cbind(Population, confirmed_US)

#Serve as data frame length and width variables
USrowmax <- nrow(confirmed_US[,]) #limit Qty of Rows 3342
UScolmax <- ncol(confirmed_US[,]) #limit Qty of Cols 1155
GloRowmax <- nrow(confirmed_global[,]) #limit Qty of Rows 289
GloColmax <- ncol(confirmed_global[,]) #limit Qty of Cols 1147

#loads column headers in data.frame
GNameCol <- as.data.frame(colnames(confirmed_global))
DNameCol <- as.data.frame(colnames(deaths_global))

```


## Instructions

*Before beginning your objectives in your final document, please state which day you downloaded the data sets on for analysis. The objectives for this lab will cumulatively cover many subjects discussed in this course and will also contain an objective for manipulating strings.*
*The surgeon general for the United States recently created a new data science initiative, CSIT-165, that uses data science to characterize pandemic diseases. CSIT-165 disseminates data driven analyses to state governors. You are a data scientist for CSIT-165 and it is up to you and you alone to manipulate and visualize COVID-19 data for disease control.*


## Objectives
*This project will encompass many of the lessons we have learned throughout the course, including interactive visualizations. RMarkdown files must be written such that each time you render the document it will download the necessary data sets for analysis. Please render the RMarkdown file the day it is due to reflect the most recent data sets. With this added functionality, your code must be able to analyze the datasets regardless of the date you render your document. Unlike others projects in the past, you will have the ability to solve these problems using any method you choose. Be careful, however, with the methods you use as you will be graded on the appropriateness of your solution and how well you execute your desired algorithm. If there appears to be a logic fail in how you executed your code, you will be penalized. This is an opportunity to really showcase your new found skills in data science with R!*


## Objective 1 *- Global Map*
*You are tasked to create a world map to gain an appreciation for where the most occurrences of COVID-19 confirmations and deaths are located.*

*Create this map using leaflet for the most recent date as shown below. For this map, sum the confirmations and deaths of provinces into one value to depict the total number for the country they belong to. When creating a marker for each country in the map, calculate lat and long as the mean values for the provinces that make up each country.*

*Customize the map to reflect the differences in magnitude for confirmations and deaths. In the example map below, circle markers that are blue represent low values, gray represents neutral values, and red represents high values. Low, middle, and high values were categorized to aesthetically map the markers based on their probabilistic distribution using the quartile function. You may use any method you like so that it is logical and allows visualization of value intensity. As well, customize the map to include hover labels that indicate country names and popup labels to show the value of confirmations and deaths for that country. For extra help using leaflet, consult this website along with the information provided in your textbooks.*


```{r ob1}
#Initiate Buffer variable
confirmed <- confirmed_global
deaths <- deaths_global

# Extract the most recent date column
latest_date <- tail(names(confirmed), 1)


summary_data <- confirmed %>%
    group_by(Country.Region) %>%
    summarise(Total_Confirmed = sum(!!sym(latest_date), na.rm = TRUE),
    Mean_Lat = mean(Lat, na.rm = TRUE), Mean_Long = mean(Long, na.rm = TRUE)) %>%
    left_join(deaths %>%
    group_by(Country.Region) %>%
    summarise(Total_Deaths = sum(!!sym(latest_date), na.rm = TRUE)), by = "Country.Region")

summary_data <- summary_data %>%
    mutate(Confirmed_Category = cut(Total_Confirmed, breaks = quantile(Total_Confirmed, probs = c(0, 0.25, 0.75, 1), na.rm = TRUE), 
    labels = c("Low", "Medium", "High"),include.lowest = TRUE), Deaths_Category = cut(Total_Deaths, breaks = quantile(Total_Deaths, 
    probs = c(0, 0.25, 0.75, 1), na.rm = TRUE), labels = c("Low", "Medium", "High"), include.lowest = TRUE))


# Define color palette for Confirmed and Deaths
confirmed_palette <- colorFactor(c("blue", "gray", "red"), levels = c("Low", "Medium", "High"))
deaths_palette <- colorFactor(c("lightblue", "lightgray", "darkred"), levels = c("Low", "Medium", "High"))


# Create the leaflet map
map <- leaflet(summary_data) %>%
    addTiles() %>%
  
    addCircleMarkers(~Mean_Long, ~Mean_Lat, label = ~paste(Country.Region), popup = ~paste0("<strong>", Country.Region, "</strong><br>",
    "Total Confirmed: ", Total_Confirmed, "<br>", "Total Deaths: ", Total_Deaths),
    color = ~confirmed_palette(Confirmed_Category), radius = ~sqrt(Total_Confirmed) / 10, group = "Confirmed") %>%
  
    addCircleMarkers(~Mean_Long, ~Mean_Lat, label = ~paste(Country.Region), popup = ~paste0("<strong>", Country.Region, 
    "</strong><br>", "Total Confirmed: ", Total_Confirmed, "<br>", "Total Deaths: ", 
    Total_Deaths), color = ~deaths_palette(Deaths_Category), radius = ~sqrt(Total_Deaths) / 10, group = "Deaths") %>%
  
    addLayersControl(overlayGroups = c("Confirmed", "Deaths"), options = layersControlOptions(collapsed = FALSE))

#Prints map
map
```

## Objective 2 *- Narrowing Down Hot Spots*
*Seeing the global map of COVID-19 cases results in the stark realization that some countries are more affected than others. In order to narrow down your studies, create a table using kable from knitr listing the top countries as far as confirmations and deaths (sum values for provinces of the same country into one value and show the country only). Now that we are using RMarkdown to create HTML files, we have much more options for how we display our table. For reference on how to customize tables using knitr, visit this website. Consult the table below for an example of a customized table ranking cases by country. While it is not required to replicate this table exactly, it would be a fantastic challenge to show off your knitr prowess.*

```{r ob2}



```

## Objective 3 *- Zooming Into Our State*

Generate an R code Program that creates a scatter plot similar to ob3 pic a.png using the following instructions. 

*After reading the top tables, you are stunned! The US overtakes every other country in terms of COVID-19 confirmations. As such, you are concerned about the state you live in and would like to understand how COVID-19 events have shaped the trajectory of the disease. Create two scatter plots to gain a better understanding. The first scatter plot should be California’s trajectory for confirmations. The second scatter plot should show California’s top three city trajectories for confirmations. You are interested in studying how the vaccine affected the number of confirmations. The Moderna vaccine was first available as an emergency use authorized (EUA) vaccine and required two shots spaced six weeks apart. Indicate on the plots the day the second dosage was given to those that received the first dosage the day Moderna was EUA (January 29th, 2021). As a diligent scientist that knows that new COVID variants have mutations in the spike protein (the region that the vaccine was developed for), you also want to study how confirmation rates change as new variants become the dominant infectious strain. Indicate on the plots when the delta and omicron variants became the dominant strain in California (May 11th, 2021 and November 26th, 2021 respectively). In the example below, the function plot_grid from the R package cowplot was to organize the graphs into a grid to more easily compare statewide vs top city plots.*

```{r ob3}


```

## Objective 4 *- Digging Deeper*


Generate an R code Program that creates a scatter plot similar to ob4 pic a.png using the following instructions. 


*Although these plots do not tell the whole story, they are great for helping us determine where to look. Different cities may have different populations, population densities, cultural discrepancies, compliance, and city regulations to name a few. We will explore the role of population on these metrics using visualizations. Arrange two scatter plots using cowplot’s plot_grid to show the relationship between population and confirmed counts as well as death counts and confirmed counts. You will need to use a log transform on all variables to show such a relationship. Please consult the example below for an idea of what this may look like. From these graphs we can see that population greatly affects confirmations and deaths. This coincides with our plots above as the population of Los Angeles is 301% greater than San Diego’s population and 406% greater than Riverside’s population!*

```{r ob4}



```

## GitHub Log

```{bash gitlog} 
git log --pretty=format:"%nSubject: %s%nAuthor: %aN%nDate: %aD%nBody: %b"
```
